\LARGE \bf The Battle for Filter Supremacy: A Comparative Study of the Multi-State Constraint Kalman Filter and the Sliding Window Filter

### 2015-03-05 Report/project_report.tex
Spacing fix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-State Constraint Kalman Filter} \label{sec:MSCKF}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds a preset threshold.
Except where otherwise noted (i.e., state parametrization and integration method), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.

\subsection{MSCKF State Parametrization}
We evaluated the MSCKF on datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
To accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.

In our implementation, we parametrize the IMU state at time $k$ as the 13-dimensional vector
\begin{align}
    \Vec{x}_{I,k} &:= \bbm \Vec{q}_{IG,k}^T & \Vec{b}_{\boldsymbol{\omega},k}^T & \Vec{b}_{\Vec{v},k}^T & \Vec{p}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
\end{align}
where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{\boldsymbol{\omega},k}$ is the bias on the gyro measurements $\boldsymbol{\omega}_m$, $\Vec{b}_{\Vec{v},k}$ is the bias on the velocity measurements $\Vec{v}_m$, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).

@@

\subsection{IMU State Estimate Propagation}
The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by a continuous-time motion model:
\begin{align}
    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} 
        & \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \notag \\
    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
        & \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} 
\end{align}
where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$,
\begin{align*}
    \hat{\Vec{v}} &= \Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}}, & \hat{\boldsymbol{\omega}} &= \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}},
\end{align*}
\begin{align*}
    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
                                                                -\hat{\boldsymbol{\omega}}^T & 0
                                                            \ebm ,
                                                            & \text{and} &&
    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
                                            \ebm .
\end{align*}
In our implementation we propagate the motion model using a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-State Constraint Kalman Filter} \label{sec:MSCKF}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds a preset threshold.
Except where otherwise noted (i.e., \ref{sec:state_parametrization} and \ref{sec:imu_prop}), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.

\subsection{MSCKF State Parametrization} \label{sec:state_parametrization}
We evaluated the MSCKF on datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
To accommodate this alternative sensor configuration, we parametrize the IMU state at time $k$ as the 13-dimensional vector
\begin{align}
    \Vec{x}_{I,k} &:= \bbm \Vec{q}_{IG,k}^T & \Vec{b}_{\boldsymbol{\omega},k}^T & \Vec{b}_{\Vec{v},k}^T & \Vec{p}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
\end{align}
where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{\boldsymbol{\omega},k}$ is the bias on the gyro measurements $\boldsymbol{\omega}_m$, $\Vec{b}_{\Vec{v},k}$ is the bias on the velocity measurements $\Vec{v}_m$, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).

@@

\subsection{IMU State Estimate Propagation} \label{sec:imu_prop}
The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by a continuous-time motion model:
\begin{align}
    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} 
        & \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \notag \\
    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
        & \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} 
\end{align}
where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$,
\begin{align*}
    \hat{\Vec{v}} &= \Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}}, & \hat{\boldsymbol{\omega}} &= \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}},
\end{align*}
\begin{align*}
    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
                                                                -\hat{\boldsymbol{\omega}}^T & 0
                                                            \ebm ,
                                                            & \text{and} &&
    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
                                            \ebm .
\end{align*}
In our implementation we propagate the motion model using a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
